---
title: "midterm"
format: html
format: 
  html:
    embed-resources: true
editor: visual
---

# Midterm Report

## Preliminaries

NOTES FROM 3/20

-   keep only california and florida
-   survey contains chemicals and codes
    -   codes turn into words?
-   pick 3 chemicals, contrast chemical use between FL and CA
    -   pick your own measures so that you find a contrast — need to see it graphically
        -   such as \# applications, lbs / acre
-   income contrast between FL and CA strawberry growers dependent on conventional, processing, and organic
-   don't need to use data that you don't find useful
-   turn echo off... so you only get words, tables, graphs
-   make sure there is writing! doesn't have to say a whole lot
-   due 3/27??
-   include linear regression in at least one
-   look at cheatsheets for help

```{r}
#| label: load libraries
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(ggplot2)
library(patchwork)
library(scales)
```

```{r}
#| label: read data - glimpse 

strawberry <- read_csv("strawb_mar6.csv", 
                       col_names = TRUE, 
                       show_col_types = FALSE)

source("my_functions.R")
```

## Data Cleaning

```{r}
#| label: drop unnecessary columns from `strawberry`

strawb <- strawberry |> drop_one_value_col()
```

```{r}
#| label: explore strawb data


test <- strawb |> show_unique(10)
```

```{r}
#| label: split strawb into two data frames

census <- strawb |> filter(Program == "CENSUS")
survey <- strawb |> filter(Program == "SURVEY")
```

```{r}
#| label: exploring unique values

census <- census |> drop_one_value_col(prt_val = TRUE)
survey <- survey |> drop_one_value_col(prt_val = TRUE)

unique_sur <- survey |> show_unique(nrows = 10)
unique_cen <- census |> show_unique(nrows = 10)

census <- census |> select(-`State ANSI`)
survey <- survey |> select(-`State ANSI`, -`Week Ending`, -`Period`)

census <- census |> filter(State == "CALIFORNIA" | State == "FLORIDA")
survey <- survey |> filter(State == "CALIFORNIA" | State == "FLORIDA")

rm(strawberry, strawb)

```

```{r}
#| label: work on census columns

commod <- census$Commodity |> unique()

# splitting Commodity
census <- census |>
  separate_wider_delim(
    cols = Commodity,
    delim = ", ",
    names = c("Income","Net Cash Farm", "Straw"),
    too_many = "error",
    #names_sep = " ",
    too_few = "align_start"
  )

# removing redundant data in Data Item
census <- census |> 
  mutate(`Data Item` = str_replace(`Data Item`, "INCOME, NET CASH FARM, ", ""))
census <- census |> 
  mutate(`Data Item` = str_replace(`Data Item`, "STRAWBERRIES, ", ""))

unique_cen <- census |> show_unique(10)
census <- census |> drop_one_value_col(prt_val = TRUE)

# splitting Data Item
census <- census |>
  separate_wider_delim(
    cols = `Data Item`,
    delim = " - ",
    names = c("Production","Category"),
    too_many = "error",
    too_few = "align_start"
  )

# splitting Production
census <- census |>
  separate_wider_delim(
    cols = Production,
    delim = ", ",
    names = c("Production", "Marketing Channel"),
    too_many = "error",
    too_few = "align_start"
  )

# splitting Category
census <- census |>
  separate_wider_delim(
    cols = Category,
    delim = ", ",
    names = c("Category", "Metric"),
    too_many = "error",
    too_few = "align_start"
  )

# removing redundant info
census <- census |> 
  mutate(`Metric` = str_replace(Metric, "MEASURED IN ", ""))

vals = c("NET GAIN: \\(", "NET LOSS: \\(", "AREA OPERATED: \\(", "ECONOMIC CLASS: \\(", "FARM SALES: \\(", "NAICS CLASSIFICATION: \\(", "ORGANIC STATUS: \\(", "\\)")
for (i in 1:length(vals)) {
  census <- census |>
  mutate(`Domain Category` = str_replace(`Domain Category`, vals[i], ""))
}

# converting strings into numerics
census$Value <- as.numeric(gsub(",", "", census$Value))
census$`CV (%)` <- as.numeric(gsub(",", "", census$`CV (%)`))
```

```{r}
#| label: work on survey columns

# splitting Data Item
survey <- survey |> separate_wider_delim(
  cols = `Data Item`,
  delim = ' - ',
  names = c("Marketing Channel", "Category"),
  too_many = "error",
  too_few = "align_start"
)

# splitting Marketing Channel
survey <- survey |> separate_wider_delim(
  cols = `Marketing Channel`,
  delim = ', ',
  names = c("Fruit", "Marketing Channel"),
  too_many = "error",
  too_few = "align_start"
)

vals = c("(D)", "(NA)")
for (i in 1:length(vals)) {
  survey <- survey |> mutate(Value = str_replace(Value, vals[i], NA_character_)) 
}

# converting strings into numerics
survey$Value <- as.numeric(gsub(",", "", survey$Value))

# splitting Category
survey <- survey |> separate_wider_delim(
  cols = `Category`,
  delim = ', ',
  names = c("Category", "Metric"),
  too_many = "merge",
  too_few = "align_start"
)

# splitting Domain
survey <- survey |> separate_wider_delim(
  cols = Domain,
  delim = ", ",
  names = c("na", "Chemical Type"),
  too_few = "align_start"
)

survey <- survey |> shift_loc("na", "FERTILIZER", 1, 1)

# removing redundant info
survey <- survey |> 
  mutate(`Metric` = str_replace(Metric, "MEASURED IN ", ""))

vals = c("CHEMICAL, HERBICIDE: \\(", "CHEMICAL, FUNGICIDE: \\(", "CHEMICAL, INSECTICIDE: \\(", "CHEMICAL, OTHER: \\(", "FERTILIZER: \\(", "\\)")
for (i in 1:length(vals)) {
 survey <- survey |>
  mutate(`Domain Category` = str_replace(`Domain Category`, vals[i], "")) 
}
survey <- survey |> 
  mutate(`Domain Category` = str_replace(`Domain Category`, "NOT SPECIFIED", NA_character_))

# splitting Domain Category
survey <- survey |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " = ",
  names = c("Chemical", "Chemical Code"),
  too_few = "align_start"
)
```

```{r}
#| label: further determining unique values

unique_cen <- show_unique(census)
unique_sur <- show_unique(survey)

```

```{r}
#| label: assignment shorter results

census_fl <- census |> filter(State == "FLORIDA")
census_ca <- census |> filter(State == "CALIFORNIA")

survey_fl <- survey |> filter(State == "FLORIDA")
survey_ca <- survey |> filter(State == "CALIFORNIA")

rm(census, survey, unique_cen, unique_sur, test)
```

```{r}
#| label: removing unnecessary columns

census_fl <- census_fl |> drop_one_value_col(prt_val = TRUE)
census_ca <- census_ca |> drop_one_value_col(prt_val = TRUE)

survey_fl <- survey_fl |> drop_one_value_col(prt_val = TRUE)
survey_ca <- survey_ca |> drop_one_value_col(prt_val = TRUE)

```

```{r}
#| label: creating chemicals data frame

chemicals_ca <- survey_ca |> filter(na == "CHEMICAL" | is.na(na))
chemicals_fl <- survey_fl |> filter(na == "CHEMICAL" | is.na(na))

chem_ca_unq <- chemicals_ca |> show_unique()
chem_fl_unq <- chemicals_fl |> show_unique()

chemicals_ca <- chemicals_ca |> select(-na)
chemicals_fl <- chemicals_fl |> select(-na)
```

```{r}
#| label: splitting bearing and non-bearing strawberries

chemicals_ca_bearing <- chemicals_ca |> filter(`Marketing Channel` == "BEARING")
chemicals_ca_nonbearing <- chemicals_ca |> filter(is.na(`Marketing Channel`))

chemicals_fl_bearing <- chemicals_fl |> filter(`Marketing Channel` == "BEARING")
chemicals_fl_nonbearing <- chemicals_fl |> filter(is.na(`Marketing Channel`))

chemicals_ca_nonbearing <- chemicals_ca_nonbearing |> select(-`Marketing Channel`, -Value)
chemicals_fl_nonbearing <- chemicals_fl_nonbearing |> select(-`Marketing Channel`, -Value)

chemicals_ca_bearing <- chemicals_ca_bearing |> select(-`Marketing Channel`)
chemicals_fl_bearing <- chemicals_fl_bearing |> select(-`Marketing Channel`)
```

```{r}
#| label: determining most used chemicals by year and state

# 2021 most used chemicals by state
most_chems_ca_21 <- chemicals_ca_bearing |> filter(Year == "2021",
                                                   Category == "APPLICATIONS",
                                                   Metric == "LB",
                                                   Chemical != "TOTAL") |>
                                            arrange(desc(Value))
most_chems_fl_21 <- chemicals_fl_bearing |> filter(Year == "2021",
                                                   Category == "APPLICATIONS",
                                                   Metric == "LB",
                                                   Chemical != "TOTAL") |>
                                            arrange(desc(Value))

# 2023 most used chemicals by state
most_chems_ca_23 <- chemicals_ca_bearing |> filter(Year == "2023",
                                                   Category == "APPLICATIONS",
                                                   Metric == "LB",
                                                   Chemical != "TOTAL") |>
                                            arrange(desc(Value))
most_chems_fl_23 <- chemicals_fl_bearing |> filter(Year == "2023",
                                                   Category == "APPLICATIONS",
                                                   Metric == "LB",
                                                   Chemical != "TOTAL")|>
                                            arrange(desc(Value))

top3_chem_ca21 <- most_chems_ca_21[1:3, ]
top3_chem_fl21 <- most_chems_fl_21[1:3, ]

top3_chem_ca23 <- most_chems_ca_23[1:3, ]
top3_chem_fl23 <- most_chems_fl_23[1:3, ]

top3_chems <- rbind(top3_chem_ca21, top3_chem_ca23, top3_chem_fl21, top3_chem_fl23)
top3_chems <- top3_chems |> mutate(State = ifelse(row_number() <= 6, "CALIFORNIA", "FLORIDA"))

rm(top3_chem_ca21, top3_chem_ca23, top3_chem_fl21, top3_chem_fl23, most_chems_ca_21, most_chems_ca_23, most_chems_fl_21, most_chems_fl_23)

```

```{r}
#| label: visualize top 3 chemical use by state

top3_21 <- top3_chems |> filter(Year == "2021")
top3_23 <- top3_chems |> filter(Year == "2023")

p1 <- ggplot(top3_21, aes(Chemical, Value, fill= `State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free_y") +
  geom_text(aes(label = Value), 
            vjust = -0.5, 
            color = "black", 
            size = 2.5) +
  scale_y_continuous(labels = label_comma()) +
  labs(
    x = "Chemical",
    y = "Value"
  ) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))

p2 <- ggplot(top3_23, aes(Chemical, Value, fill= `State`)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free_y") +
  geom_text(aes(label = Value), 
            vjust = -0.5, 
            color = "black", 
            size = 2.5) + 
  scale_y_continuous(labels = label_comma()) +
  labs(
    x = "Chemical",
    y = "Value"
  ) + 
  theme_minimal() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1), plot.title = element_text(hjust = 0.5))


(p1 + p2) +
  plot_layout(guides = "collect") +
  plot_annotation(title = "Most Used Chemicals by State and Year") &
  theme(legend.position = "bottom",
        plot.title = element_text(hjust = 0.5, size = 14, face = "bold"))
```

```{r}
#| label: determining common chemicals used by both FL and CA

chemicals_fl <- chemicals_fl |> mutate(State = "FLORIDA")
chemicals_ca <- chemicals_ca |> mutate(State = "CALIFORNIA")
common_chems <- rbind(chemicals_fl, chemicals_ca)
common_chems <- common_chems |> drop_na(Value)

```

Based on the graph above and the `common_chems` data frame, I will explore Thiram, Captan, and Fludioxonil.

## Exploring Chemicals

```{r}

```
