---
title: "Midterm Report"
format: 
  html:
    embed-resources: true
editor: visual
execute:
  echo: false
---

I am starting with cleaning the data and then separating it into 4 data sets: California census, California survey, Florida census, and Florida survey.

```{r}
#| label: load libraries
#| warning: false
#| message: false

library(knitr)  
library(kableExtra)
library(tidyverse)
library(stringr)
library(ggplot2)
library(patchwork)
library(scales)
library(readr)
```

```{r}
#| label: read data - glimpse 

strawberry <- read_csv("strawb_mar6.csv", 
                       col_names = TRUE, 
                       show_col_types = FALSE)

source("my_functions.R")
```

```{r}
#| label: drop unnecessary columns from `strawberry`

strawb <- strawberry |> drop_one_value_col()
```

```{r}
#| label: explore strawb data
#| warning: false


test <- strawb |> show_unique(10)
```

```{r}
#| label: split strawb into two data frames

census <- strawb |> filter(Program == "CENSUS")
survey <- strawb |> filter(Program == "SURVEY")
```

```{r}
#| label: exploring unique values

census <- census |> drop_one_value_col()
survey <- survey |> drop_one_value_col()

unique_sur <- survey |> show_unique(nrows = 10)
unique_cen <- census |> show_unique(nrows = 10)

census <- census |> select(-`State ANSI`)
survey <- survey |> select(-`State ANSI`, -`Week Ending`, -`Period`)

census <- census |> filter(State == "CALIFORNIA" | State == "FLORIDA")
survey <- survey |> filter(State == "CALIFORNIA" | State == "FLORIDA")

rm(strawberry, strawb)

```

```{r}
#| label: work on census columns
#| warning: false

commod <- census$Commodity |> unique()

# splitting Commodity
census <- census |>
  separate_wider_delim(
    cols = Commodity,
    delim = ", ",
    names = c("Income","Net Cash Farm", "Straw"),
    too_many = "error",
    #names_sep = " ",
    too_few = "align_start"
  )

# removing redundant data in Data Item
census <- census |> 
  mutate(`Data Item` = str_replace(`Data Item`, "INCOME, NET CASH FARM, ", ""))
census <- census |> 
  mutate(`Data Item` = str_replace(`Data Item`, "STRAWBERRIES, ", ""))

unique_cen <- census |> show_unique(10)
census <- census |> drop_one_value_col()

# splitting Data Item
census <- census |>
  separate_wider_delim(
    cols = `Data Item`,
    delim = " - ",
    names = c("Production","Category"),
    too_many = "error",
    too_few = "align_start"
  )

census <- census |> mutate(Production = str_replace(Production, "OF ", ""))

# splitting Production
census <- census |>
  separate_wider_delim(
    cols = Production,
    delim = ", ",
    names = c("Production", "Marketing Channel"),
    too_many = "error",
    too_few = "align_start"
  )


vals = c("NET GAIN: \\(", "NET LOSS: \\(", "AREA OPERATED: \\(", "ECONOMIC CLASS: \\(", "FARM SALES: \\(", "NAICS CLASSIFICATION: \\(", "ORGANIC STATUS: \\(", "\\)")
for (i in 1:length(vals)) {
  census <- census |>
  mutate(`Domain Category` = str_replace(`Domain Category`, vals[i], ""))
}

# converting strings into numerics
census$Value <- parse_number(census$Value)
census <- census |> mutate(`CV (%)` = str_replace(`CV (%)`, "(H)", "99.95"))
census$`CV (%)` <- parse_number(census$`CV (%)`)
```

```{r}
#| label: work on survey columns
#| warning: false

# splitting Data Item
survey <- survey |> separate_wider_delim(
  cols = `Data Item`,
  delim = ' - ',
  names = c("Marketing Channel", "Category"),
  too_many = "error",
  too_few = "align_start"
)

# splitting Marketing Channel
survey <- survey |> separate_wider_delim(
  cols = `Marketing Channel`,
  delim = ', ',
  names = c("Fruit", "Marketing Channel"),
  too_many = "error",
  too_few = "align_start"
)

# converting strings into numerics
vals = c("(D)", "(NA)")
for (i in 1:length(vals)) {
  survey <- survey |> mutate(Value = str_replace(Value, vals[i], NA_character_)) 
}
survey$Value <- parse_number(survey$Value)

# splitting Category
survey <- survey |> separate_wider_delim(
  cols = Category,
  delim = ", MEASURED IN ",
  names = c("Category", "Metric"),
  too_few = "align_start"
)

survey <- survey |> mutate(Category = str_replace(Category, "ACRES HARVESTED", "AREA HARVESTED"))

survey$Metric[survey$Category == "AREA HARVESTED" & is.na(survey$Metric)] <- "ACRES"

# splitting Domain
survey <- survey |> separate_wider_delim(
  cols = Domain,
  delim = ", ",
  names = c("na", "Chemical Type"),
  too_few = "align_start"
)

survey <- survey |> shift_loc("na", "FERTILIZER", 1, 1)


vals = c("CHEMICAL, HERBICIDE: \\(", "CHEMICAL, FUNGICIDE: \\(", "CHEMICAL, INSECTICIDE: \\(", "CHEMICAL, OTHER: \\(", "FERTILIZER: \\(", "\\)")
for (i in 1:length(vals)) {
 survey <- survey |>
  mutate(`Domain Category` = str_replace(`Domain Category`, vals[i], "")) 
}
survey <- survey |> 
  mutate(`Domain Category` = str_replace(`Domain Category`, "NOT SPECIFIED", NA_character_))

# splitting Domain Category
survey <- survey |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " = ",
  names = c("Chemical", "Chemical Code"),
  too_few = "align_start"
)

rm(vals, i, commod)
```

```{r}
#| label: assignment shorter results

## creating census and survey datasets for both Florida and California

census_fl <- census |> filter(State == "FLORIDA")
census_ca <- census |> filter(State == "CALIFORNIA")

survey_fl <- survey |> filter(State == "FLORIDA")
survey_ca <- survey |> filter(State == "CALIFORNIA")

rm(census, survey, unique_cen, unique_sur, test)
```

```{r}
#| label: removing unnecessary columns

census_fl <- census_fl |> drop_one_value_col()
census_ca <- census_ca |> drop_one_value_col()

survey_fl <- survey_fl |> drop_one_value_col()
survey_ca <- survey_ca |> drop_one_value_col()

```

# California

All the data cleaning for California is done in this section.

## Census Data

In this section, I'm cleaning the census data for California. This is where I am getting the information for the operations vs producers section.

```{r}
#| label: splitting census_ca into 2021 and 2022

census_ca_22 <- census_ca |> filter(Year == 2022)
census_ca_21 <- census_ca |> filter(Year == 2021)

census_ca_22 <- census_ca_22 |> drop_one_value_col()
census_ca_21 <- census_ca_21 |> drop_one_value_col()

```

```{r}
#| label: splitting census_ca_22 into producers and operators

census_ca_22_pr <- census_ca_22 |> filter(Production == "PRODUCERS")
census_ca_22_op <- census_ca_22 |> filter(Production == "OPERATIONS")

census_ca_22_pr <- census_ca_22_pr |> drop_one_value_col()
census_ca_22_op <- census_ca_22_op |> drop_one_value_col()


```

```{r}
#| label: exploring operations vs producers, net income and area operated in California
#| warning: false

net_inc_ca_pr <- census_ca_22_pr |> filter(Category == "NET INCOME, MEASURED IN $" & Domain == "AREA OPERATED")
net_inc_ca_op <- census_ca_22_op |> filter(Category == "NET INCOME, MEASURED IN $" & Domain == "AREA OPERATED")

net_inc_ca_pr <- net_inc_ca_pr |> mutate(`Domain Category` = str_replace(`Domain Category`, " ACRES", ""))
net_inc_ca_op <- net_inc_ca_op |> mutate(`Domain Category` = str_replace(`Domain Category`, " ACRES", ""))
net_inc_ca_pr <- net_inc_ca_pr |> mutate(`Domain Category` = str_replace(`Domain Category`, "OR", "TO"))
net_inc_ca_op <- net_inc_ca_op |> mutate(`Domain Category` = str_replace(`Domain Category`, "OR", "TO"))

net_inc_ca_pr <- net_inc_ca_pr |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " TO ",
  names = c("Low", "High"),
  too_few = "align_start",
  too_many = "error"
)

net_inc_ca_op <- net_inc_ca_op |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " TO ",
  names = c("Low", "High"),
  too_few = "align_start",
  too_many = "error"
)

net_inc_ca_pr$Low <- parse_number(net_inc_ca_pr$Low)
net_inc_ca_pr$High <- parse_number(net_inc_ca_pr$High)
net_inc_ca_op$Low <- parse_number(net_inc_ca_op$Low)
net_inc_ca_op$High <- parse_number(net_inc_ca_op$High)

net_inc_ca_pr <- net_inc_ca_pr |> arrange(Low)
net_inc_ca_op <- net_inc_ca_op |> arrange(Low)

net_inc_ca_pr <- net_inc_ca_pr |> rename("Value ($)" = Value)
net_inc_ca_op <- net_inc_ca_op |> rename("Value ($)" = Value)
```

```{r}
#| label: creating a table comparing net income to area operated in California

net_inc_ca_pr_table <- net_inc_ca_pr |> drop_one_value_col()
net_inc_ca_op_table <- net_inc_ca_op |> drop_one_value_col()

table1 <- net_inc_ca_pr_table |> kable()
table2 <- net_inc_ca_op_table |> kable()

table1 <- net_inc_ca_pr_table |> kbl(caption = "California's Producers Net Income in 2022 by Area Operated") |> kable_classic(full_width = F, html_font = "Cambria") |> add_header_above(c( "Area Range (acres)" = 2, " " = 1, " " = 1))

table2 <- net_inc_ca_op_table |> kbl(caption = "California's Operations Net Income in 2022 by Area Operated") |> kable_classic(full_width = F, html_font = "Cambria") |> add_header_above(c( "Area Range (acres)" = 2, " " = 1, " " = 1))
```

## Survey Data

I am cleaning the survey data for California in this section. Here is where I am discovering the information I use for California's chemical analysis.

```{r}
#| label: splitting survey_ca into years

survey_ca_23 <- survey_ca |> filter(Year == 2023)
survey_ca_22 <- survey_ca |> filter(Year == 2022)
survey_ca_21 <- survey_ca |> filter(Year == 2021)
survey_ca_20 <- survey_ca |> filter(Year == 2020)

survey_ca_23 <- survey_ca_23 |> drop_one_value_col()
survey_ca_22 <- survey_ca_22 |> drop_one_value_col()
survey_ca_21 <- survey_ca_21 |> drop_one_value_col()
survey_ca_20 <- survey_ca_20 |> drop_one_value_col()
```

```{r}
#| label: determining chemicals used in California

# only the 2021 and 2023 tibbles contain chemical information
chemicals_ca_21 <- survey_ca_21 |> filter(!is.na(Chemical) & Chemical != "TOTAL")
chemicals_ca_23 <- survey_ca_23 |> filter(!is.na(Chemical) & Chemical != "TOTAL")

chemicals_ca_21 <- chemicals_ca_21 |> drop_one_value_col()
chemicals_ca_23 <- chemicals_ca_23 |> drop_one_value_col()

# removing unnecessary details
chemicals_ca_23 <- chemicals_ca_23 |> select(-na)
chemicals_ca_23 <- chemicals_ca_23 |> filter(!(is.na(`Marketing Channel`) & is.na(Value)))
chemicals_ca_23 <- chemicals_ca_23 |> drop_one_value_col()
```

```{r}
#| label: finding the chemicals that were used the most in applications in California

chems_apps_ca_21 <- chemicals_ca_21 |> filter(Category == "APPLICATIONS" & Metric == "LB") |> arrange(desc(Value))
chems_apps_ca_21 <- chems_apps_ca_21 |> filter(!is.na(Value))

chems_apps_ca_23 <- chemicals_ca_23 |> filter(Category == "APPLICATIONS" & Metric == "LB") |> arrange(desc(Value))
chems_apps_ca_23 <- chems_apps_ca_23 |> filter(!is.na(Value))

top5_chem_apps_ca21 <- chems_apps_ca_21[1:5,]
top5_chem_apps_ca23 <- chems_apps_ca_23[1:5,]

top5_chem_apps_ca21$Year <- "2021"
top5_chem_apps_ca23$Year <- "2023"
top5_chems_ca_combined <- rbind(top5_chem_apps_ca21, top5_chem_apps_ca23)

plot1_top <- ggplot(top5_chems_ca_combined, aes(x=Chemical, y=Value, fill=Year)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free") +
  geom_text(aes(label = Value), vjust = -0.5, size = 2.5) +
  labs(
    title = "Top 5 Chemicals Used in California by Year",
    x = "Chemical",
    y = "Applications (in lbs)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

bottom5_chem_apps_ca21 <- tail(chems_apps_ca_21, 5)
bottom5_chem_apps_ca23 <- tail(chems_apps_ca_23, 5)

bottom5_chem_apps_ca21$Year <- "2021"
bottom5_chem_apps_ca23$Year <- "2023"
bottom5_chems_ca_combined <- rbind(bottom5_chem_apps_ca21, bottom5_chem_apps_ca23)

plot1_bottom <- ggplot(bottom5_chems_ca_combined, aes(x=Chemical, y=Value, fill=Year)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free") +
  geom_text(aes(label = Value), vjust = -0.5, size = 2.5) +
  labs(
    title = "Bottom 5 Chemicals Used in California by Year",
    x = "Chemical",
    y = "Applications (in lbs)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
```

# Florida

All the data cleaning for Florida is done in this section.

## Census Data

In this section, I'm cleaning the census data for Florida. Similarly to California, this is where I am getting the information for the operations vs producers section.

```{r}
#| label: splitting census_fl into 2021 and 2022

census_fl_22 <- census_fl |> filter(Year == 2022)
census_fl_21 <- census_fl |> filter(Year == 2021)

census_fl_22 <- census_fl_22 |> drop_one_value_col()
census_fl_21 <- census_fl_21 |> drop_one_value_col()

```

```{r}
#| label: splitting census_fl_22 into producers and operators

census_fl_22_pr <- census_fl_22 |> filter(Production == "PRODUCERS")
census_fl_22_op <- census_fl_22 |> filter(Production == "OPERATIONS")

census_fl_22_pr <- census_fl_22_pr |> drop_one_value_col()
census_fl_22_op <- census_fl_22_op |> drop_one_value_col()


```

```{r}
#| label: exploring operations vs producers, net income and area operated in Florida
#| warning: false

net_inc_fl_pr <- census_fl_22_pr |> filter(Category == "NET INCOME, MEASURED IN $" & Domain == "AREA OPERATED")
net_inc_fl_op <- census_fl_22_op |> filter(Category == "NET INCOME, MEASURED IN $" & Domain == "AREA OPERATED")

net_inc_fl_pr <- net_inc_fl_pr |> mutate(`Domain Category` = str_replace(`Domain Category`, " ACRES", ""))
net_inc_fl_op <- net_inc_fl_op |> mutate(`Domain Category` = str_replace(`Domain Category`, " ACRES", ""))
net_inc_fl_pr <- net_inc_fl_pr |> mutate(`Domain Category` = str_replace(`Domain Category`, "OR", "TO"))
net_inc_fl_op <- net_inc_fl_op |> mutate(`Domain Category` = str_replace(`Domain Category`, "OR", "TO"))

net_inc_fl_pr <- net_inc_fl_pr |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " TO ",
  names = c("Low", "High"),
  too_few = "align_start",
  too_many = "error"
)

net_inc_fl_op <- net_inc_fl_op |> separate_wider_delim(
  cols = `Domain Category`,
  delim = " TO ",
  names = c("Low", "High"),
  too_few = "align_start",
  too_many = "error"
)

net_inc_fl_pr$Low <- parse_number(net_inc_fl_pr$Low)
net_inc_fl_pr$High <- parse_number(net_inc_fl_pr$High)
net_inc_fl_op$Low <- parse_number(net_inc_fl_op$Low)
net_inc_fl_op$High <- parse_number(net_inc_fl_op$High)

net_inc_fl_pr <- net_inc_fl_pr |> arrange(Low)
net_inc_fl_op <- net_inc_fl_op |> arrange(Low)

net_inc_fl_pr <- net_inc_fl_pr |> rename("Value ($)" = Value)
net_inc_fl_op <- net_inc_fl_op |> rename("Value ($)" = Value)
```

```{r}
#| label: creating a table comparing net income to area operated in Florida

net_inc_fl_pr_table <- net_inc_fl_pr |> drop_one_value_col()
net_inc_fl_op_table <- net_inc_fl_op |> drop_one_value_col()

table3 <- net_inc_fl_pr_table |> kable()
table4 <- net_inc_fl_op_table |> kable()

table3 <- net_inc_fl_pr_table |> kbl(caption = "Florida's Producers Net Income in 2022 by Area Operated") |> kable_classic(full_width = F, html_font = "Cambria") |> add_header_above(c( "Area Range (acres)" = 2, " " = 1, " " = 1))

table4 <- net_inc_fl_op_table |> kbl(caption = "Florida's Operations Net Income in 2022 by Area Operated") |> kable_classic(full_width = F, html_font = "Cambria") |> add_header_above(c( "Area Range (acres)" = 2, " " = 1, " " = 1))
```

## Survey Data

I am cleaning the survey data for Florida in this section. Like with California, here is where I am discovering the information I use for Florida chemical analysis.

```{r}
#| label: splitting survey_fl into years

survey_fl_23 <- survey_fl |> filter(Year == 2023)
survey_fl_22 <- survey_fl |> filter(Year == 2022)
survey_fl_21 <- survey_fl |> filter(Year == 2021)
survey_fl_20 <- survey_fl |> filter(Year == 2020)

survey_fl_23 <- survey_fl_23 |> drop_one_value_col()
survey_fl_22 <- survey_fl_22 |> drop_one_value_col()
survey_fl_21 <- survey_fl_21 |> drop_one_value_col()
survey_fl_20 <- survey_fl_20 |> drop_one_value_col()
```

```{r}
#| label: determining chemicals used in Florida

# only the 2021 and 2023 tibbles contain chemical information
chemicals_fl_21 <- survey_fl_21 |> filter(!is.na(Chemical) & Chemical != "TOTAL")
chemicals_fl_23 <- survey_fl_23 |> filter(!is.na(Chemical) & Chemical != "TOTAL")

chemicals_fl_21 <- chemicals_fl_21 |> drop_one_value_col()
chemicals_fl_23 <- chemicals_fl_23 |> drop_one_value_col()

# removing unnecessary details
chemicals_fl_23 <- chemicals_fl_23 |> select(-na)
chemicals_fl_23 <- chemicals_fl_23 |> filter(!(is.na(`Marketing Channel`) & is.na(Value)))
chemicals_fl_23 <- chemicals_fl_23 |> drop_one_value_col()
```

```{r}
#| label: finding the chemicals that were used the most in applications in Florida

chems_apps_fl_21 <- chemicals_fl_21 |> filter(Category == "APPLICATIONS" & Metric == "LB") |> arrange(desc(Value))
chems_apps_fl_21 <- chems_apps_fl_21 |> filter(!is.na(Value))

chems_apps_fl_23 <- chemicals_fl_23 |> filter(Category == "APPLICATIONS" & Metric == "LB") |> arrange(desc(Value))
chems_apps_fl_23 <- chems_apps_fl_23 |> filter(!is.na(Value))

top5_chem_apps_fl21 <- chems_apps_fl_21[1:5,]
top5_chem_apps_fl23 <- chems_apps_fl_23[1:5,]

top5_chem_apps_fl21$Year <- "2021"
top5_chem_apps_fl23$Year <- "2023"
top5_chems_fl_combined <- rbind(top5_chem_apps_fl21, top5_chem_apps_fl23)

plot2_top <- ggplot(top5_chems_fl_combined, aes(x=Chemical, y=Value, fill=Year)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free") +
  geom_text(aes(label = Value), vjust = -0.5, size = 2.5) +
  labs(
    title = "Top 5 Chemicals Used in Florida by Year",
    x = "Chemical",
    y = "Applications (in lbs)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

bottom5_chem_apps_fl21 <- tail(chems_apps_fl_21, 5)
bottom5_chem_apps_fl23 <- tail(chems_apps_fl_23, 5)

bottom5_chem_apps_fl21$Year <- "2021"
bottom5_chem_apps_fl23$Year <- "2023"
bottom5_chems_fl_combined <- rbind(bottom5_chem_apps_fl21, bottom5_chem_apps_fl23)

plot2_bottom <- ggplot(bottom5_chems_fl_combined, aes(x=Chemical, y=Value, fill=Year)) + 
  geom_bar(stat = "identity") +
  facet_wrap(~ Year, scales = "free") +
  geom_text(aes(label = Value), vjust = -0.5, size = 2.5) +
  labs(
    title = "Bottom 5 Chemicals Used in Florida by Year",
    x = "Chemical",
    y = "Applications (in lbs)"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1),
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14))
```

# Chemical Analysis

First, I'm comparing the top 5 chemicals used in each state.

```{r}
#| label: comparing top 5 chemicals in each state

plot1_top
plot2_top

## shared chemicals between the top 5 chemicals in each state include: CAPTAN and NITROGEN
```

Now I'm comparing the the bottom 5 chemicals used in each state.

```{r}
#| label: comparing bottom 5 chemicals in each state

plot1_bottom
plot2_bottom

## shared chemicals between the bottom 5 chemicals in each state include: ABAMECTIN and THIAMETHOXAM
```

Based on the above graphs, shared chemicals between the top 5 chemicals in each state include captan and nitrogen, and the bottom 5 chemicals in each state include abamectin and thiamethoxam. Using this information, I will compare the usage of captan, nitrogen, and abamectin between California and Florida.

```{r}
#| label: creating tables of Captan, Nitrogen, and Abamectin

top5_chems_ca_combined$State = "CALIFORNIA"
bottom5_chems_ca_combined$State = "CALIFORNIA"
top5_chems_fl_combined$State = "FLORIDA"
bottom5_chems_fl_combined$State = "FLORIDA"
chemicals_combined <- rbind(top5_chems_ca_combined, bottom5_chems_ca_combined, top5_chems_fl_combined, bottom5_chems_fl_combined)

# removing some variables since not needed
rm(top5_chems_ca_combined, bottom5_chems_ca_combined, top5_chems_fl_combined, bottom5_chems_fl_combined, top5_chem_apps_ca21, top5_chem_apps_ca23, top5_chem_apps_fl21, top5_chem_apps_fl23, bottom5_chem_apps_ca21, bottom5_chem_apps_ca23, bottom5_chem_apps_fl21, bottom5_chem_apps_fl23, plot1_bottom, plot1_top, plot2_bottom, plot2_top, chemicals_ca_21, chemicals_ca_23, chemicals_fl_21, chemicals_fl_23, chems_apps_ca_21, chems_apps_ca_23, chems_apps_fl_21, chems_apps_fl_23)

chemicals_combined <- chemicals_combined |> filter(Chemical == "CAPTAN" | Chemical == "NITROGEN" | Chemical == "ABAMECTIN")

chemicals_combined <- unique(chemicals_combined)

captan_usage <- chemicals_combined |> filter(Chemical == "CAPTAN") 
captan_usage <- captan_usage |> drop_one_value_col()

nitrogen_usage <- chemicals_combined |> filter(Chemical == "NITROGEN") 
nitrogen_usage <- nitrogen_usage |> drop_one_value_col()

abamectin_usage <- chemicals_combined |> filter(Chemical == "ABAMECTIN") 
abamectin_usage <- abamectin_usage |> drop_one_value_col()
abamectin_usage <- add_row(abamectin_usage, Value = NA, Year = "2021", State = "FLORIDA")
abamectin_usage <- abamectin_usage |> mutate(Label = as.character(Value))
abamectin_usage[4, 1] <- 0
abamectin_usage[4, 4] <- "NA"
```

```{r}
#| label: comparing usage of Captan, Nitrogen, and Abamectin

captan_plot <- ggplot(captan_usage, aes(Year, Value, fill=State)) +
  geom_bar(stat = "identity") + 
  facet_wrap(~ State) +
  geom_text(aes(label = Value), vjust = -0.5, size = 3) +
  labs(
    title = "Captan Usage by State and Year",
    x = "Year",
    y = "Applications (in lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

nitrogen_plot <- ggplot(nitrogen_usage, aes(State, Value, fill=State)) +
  geom_bar(stat = "identity") + 
  geom_text(aes(label = Value), vjust = -0.5, size = 3) +
  labs(
    title = "Nitrogen Usage by State in 2023",
    x = "State",
    y = "Applications (in lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

abamectin_plot <- ggplot(abamectin_usage, aes(Year, Value, fill=State)) +
  geom_bar(stat = "identity") + 
  facet_wrap(~ State) +
  geom_text(aes(label = Label), vjust = -0.5, size = 3) +
  labs(
    title = "Abamectin Usage by State and Year",
    x = "Year",
    y = "Applications (in lbs)"
  ) +
  theme(plot.title = element_text(hjust = 0.5, face = "bold", size = 14))

nitrogen_plot
abamectin_plot
captan_plot

```

Nitrogen was only used in 2023 in both California and Florida. There is a 110,000 lb difference in application between states. I wonder if this is due to California producing more strawberries than Florida. Also, why is there no nitrogen usage in 2021 for both states?

In regard to abamectin, both California and Florida used more in 2023 than 2021. Although Florida in 2021 doesn't have any data, I will assume that there was little to no usage. If I assume there is no usage, then both California and Florida use 100 lbs more in 2023 than in 2021.

However, the difference in captan usage between states is drastic. There is a 349,500 lb difference between 2021 and 2023 in California and a 8,900 lb difference between 2021 and 2023 in Florida. What happened between 2021 and 2023 in California for the dramatic increase in captan usage? And why is there such a small difference in captan usage from 2021 to 2023 in Florida while there's a huge difference in California?

# Operations vs Producers

```{r}
#| label: comparing operations and producers in each state

# comparing operations and producers in California
comparison_table_ca <- net_inc_ca_pr_table |>
  left_join(net_inc_ca_op_table, by = c("Low", "High"))  # Adjust "Area" to your common column

comparison_table_ca <- comparison_table_ca |> select(-`CV (%).x`, -`CV (%).y`)
comparison_table_ca <- comparison_table_ca |> mutate("Difference in y-x ($) " = `Value ($).y` - `Value ($).x`)

# comparing operations and producers in Florida
comparison_table_fl <- net_inc_fl_pr_table |>
  left_join(net_inc_fl_op_table, by = c("Low", "High"))  # Adjust "Area" to your common column

comparison_table_fl <- comparison_table_fl |> select(-`CV (%).x`, -`CV (%).y`)
comparison_table_fl <- comparison_table_fl |> mutate("Difference in y-x ($) " = `Value ($).y` - `Value ($).x`)
```

```{r}
#| label: producing the tables for comparison

# table for California
comparison_table_ca |>
  kbl(caption = "Comparison of California's Producers and Operations Net Income in 2022 by Area Operated") |>
  kable_classic(full_width = FALSE, html_font = "Cambria") |>
  add_header_above(c("Area Range (acres)" = 2, "Producers" = 1, "Operations" = 1, "Difference" = 1))

# table for Florida
comparison_table_fl |>
  kbl(caption = "Comparison of Florida's Producers and Operations Net Income in 2022 by Area Operated") |>
  kable_classic(full_width = FALSE, html_font = "Cambria") |>
  add_header_above(c("Area Range (acres)" = 2, "Producers" = 1, "Operations" = 1, "Difference" = 1))
```

Based on the tables above, we see that the net income of operations is significantly higher than the net income of producers for both California and Florida. This suggests that larger agricultural businesses (operations) are more profitable due to factors such as better market access. However, producers likely face higher costs, which leads to a lower net income.

We also see that California tends to see a higher profit than Florida in most of the area ranges. This suggests that California grows more strawberries than Florida, which can be further validated by this data set.

# Further Analyses

There were many topics explored above in regard to chemical analyses and operations/producers analyses. However, there is always more exploration to be done in a data set as large as this one. I suggest further analyses into the chemicals, specifically why some chemicals are used in one state and not in the other. Additionally, I'm curious to know what happened between 2021 and 2023 for there to be such an increase in chemical usage. On top of that, it would be good to know why there is no chemical information in 2020 and 2022.

Regarding the operations/producers analyses, it would be interesting to learn the specifics behind why there is such a large difference in profit between the states. Is it due to only one reason, such as California producing more strawberries than Florida, or are there other reasons that have yet to be explored or seen in this data set?

There are plenty more questions to be answered from this strawberry data set. I suggest the further exploration of the analyses above, in addition to other questions that an analyst may have.
